---
title: "Project №1 Burankova"
output: 
  html_document:
    df_print: paged
    toc: true
    toc_float:
      collapsed: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(scipen = 1, digits = 6)

```

# Import and data checking

## Import required library
**tidyverse** library is used for processing data.\

```{r message=FALSE, warning=FALSE}
library(tidyverse)
```

This library include all required libraries:\
- **ggplot2**, for data visualization.\
- **dplyr**, for data manipulation.\
- **tidyr**, for data tidying.\
- **readr**, for data import.\
- **purrr**, for functional programming.\

tidyverse 1.3.1:
v ggplot2 3.3.3     v purrr   0.3.4
v tibble  3.1.0     v dplyr   1.0.5
v tidyr   1.1.3     v stringr 1.4.0
v readr   1.4.0     v forcats 0.5.1


## Import .csv data
For import all .csv file to one list, run:
```{r message=FALSE, warning=FALSE}
all_participants <-
  list.files(path = "../data/", pattern = "*.csv", full.names = TRUE) %>% 
  map_df(~read_csv(.))
```
.csv files with raw data should be placed in /data/ folder.\
The data:
```{r echo=FALSE}
head(all_participants)
```


## Check and fix the data

For each column, we check for wrong values:\
- if we know the missing value for the participant (by ID) from other rows from the same year - we change it.\
- if we don`t know anything - change to NA.\
We do not delete any records, even if they contain only information about the participant ID about which there are more complete records. Removing such "truncated" rows would distort information like "the number of participation at the each Olympiad for each participant".\
Thus, we will save the maximum information about each of the participants.


### "Sex" column
Check "Sex" column.
```{r echo=FALSE}
all_participants %>% filter(!(Sex %in% c('F', 'M', NA)))

```

Fix wrong values to correct (if known - for ID == '79609') or to NA:
```{r}
#Check if we know something about the participants:
all_participants %>% filter(ID == '79609' | ID == '79630')

all_participants <- all_participants %>% mutate(Sex = ifelse(ID == '79609', 'M', Sex))
all_participants <- all_participants %>% 
  mutate(Sex = ifelse(!Sex %in% c('F', 'M', NA), 'NA', Sex))
```

### "Sport" column

```{r echo=FALSE}
all_participants %>% group_by(Sport) %>% summarise()
```

After reviewing the data manually, there find some problems with "football":

```{r echo=FALSE}
all_participants %>% filter(str_detect(Sport, "Footba")) %>% group_by(Sport) %>% 
  summarise()
```
Change 'footba' to football:
```{r}
all_participants <- all_participants %>% 
  mutate(Sport = ifelse(Sport == "Footba", 'Football', Sport))
```

### "Name" column
Let's check the correctness of the participants names.\
Each participant is assigned an ID. The ID is assigned sequentially, one participant for each ID.\
The number of participants names have to match the number of IDs.

```{r message=FALSE, warning=FALSE, include=FALSE}
ID <- all_participants %>% group_by(ID, Name) %>% summarise() %>% ungroup() %>% 
  mutate(RowNumber = 1:n()) %>% tail(1)
```
Do we have duplicate participants and how many are there: \
- 'true' if there are no duplicates: **`r ID[3] - ID[1] == 0`**.\
- how many duplicates: **`r (ID[3] - ID[1])[1,1]`**.\

In this case only one participant register twice.\
Find the participant registered twice under different names: ID - `r format(all_participants %>% group_by(ID, Name) %>% summarise() %>% ungroup() %>% mutate(RowNumber = 1:n()) %>% filter(RowNumber != ID) %>% head(1)[1,1], scientific = FALSE)`.

```{r echo=FALSE, message=FALSE, warning=FALSE}

wrong_ID <- (all_participants %>% 
               group_by(ID, Name) %>%
               summarise() %>% 
               ungroup() %>% 
               mutate(RowNumber = 1:n()) %>%
               filter(RowNumber != ID) %>%  
               head(1)[1,1])


all_participants %>% filter(ID == wrong_ID)

```

Let's fix it:

```{r}

all_participants <- all_participants %>% 
  mutate(Name = ifelse((Name == 'Pietro Spec' & ID == '113716'), 'Pietro Speciale', Name))

```

If we have a lot of wrong names and it is impossible to manually fix it by setting the correct name - assign the names to constant number.\
But in this case, we lose information about names for duplicate participants. Therefore, one should manually change the names whenever possible.

```{r message=FALSE, warning=FALSE}

f_names <- function(all_participants) {
  all_participants %>% group_by(ID, Name) %>% summarise() %>% ungroup() %>% 
    mutate(RowNumber = 1:n()) %>% filter(RowNumber != ID) %>% head(1)[1,1,1]
}

for (i in c(1:(ID[3] - ID[1])[1,1])){
  all_participants <- all_participants %>% 
    mutate(Name = ifelse((ID == f_names(all_participants)) & !is.na(f_names(all_participants)),
                         toString(i), Name))
}

```


### "Height" column
The tallest person in the world was 272 cm tall. Let's check our data:
```{r echo=FALSE}
all_participants %>% filter(ID %in% (all_participants %>% 
                                     filter(Height > 272) %>%  
                                     .[,1,1]) &
                            Age %in% (all_participants %>% 
                                      filter(Height > 272) %>%  
                                      .[,4,1]))

```

As can be seen from data, the participant #23549 is 176 cm tall. Let's fix it:
```{r}
all_participants <- all_participants %>% 
  mutate(Height = ifelse((ID == '23549' & Height > 272), 176, Height))
```


If it is impossible to manually fix it - assign the height to NA:
```{r}

all_participants <- all_participants %>% 
  mutate(Height = ifelse(Height > 272, NA, Height))

```


### 'Age' column
The oldest person in the world died aged 122.
```{r echo=FALSE}
all_participants %>% filter(ID %in% (all_participants %>% 
                                    filter(Age > 122) %>%  
                                    .[,1,1]) &
                            Year %in% (all_participants %>% 
                                      filter(Age > 122) %>%  
                                      .[,10,1]))
```
In 1912 he was 24. Fix:
```{r}
all_participants <- all_participants %>% 
  mutate(Age = ifelse((ID == '23459' & Age == 240), 24, Age))
```

If it is impossible to manually fix it - assign the age to NA:
```{r}
all_participants <- all_participants %>% 
  mutate(Age = ifelse(Age > 122, NA, Age))
```

The youngest participant in the Games was about 8 years old - it is a record.\
Let's check if there are younger members in the data. If the record was not broken - replace incorrect value manually or to NA

```{r echo=FALSE}
all_participants %>% filter(ID %in% (all_participants %>% 
                                    filter(Age < 8) %>%  
                                    .[,1,1]) &
                            Year %in% (all_participants %>% 
                                      filter(Age < 8) %>%  
                                      .[,10,1]))
```


### 'Weight' column
According to WHO, for a person 8 years old, a weight of 17 kg is considered very low. Let's check if we have participants with extremely low weight. 

```{r echo=FALSE}
all_participants %>% filter(ID %in% (all_participants %>% 
                                     filter(Weight < 17) %>%  
                                     .[,1,1]))
```
At 25 he was 77 kg. Fix:
```{r}
all_participants <- all_participants %>% 
  mutate(Weight = ifelse((ID == '68370' & Weight == 7), 77, Weight))
```

If it is impossible to manually fix it - assign the age to NA:
```{r}
all_participants <- all_participants %>% 
  mutate(Weight = ifelse(Weight < 17, NA, Weight))
```

The fattest man in the world weighed 445 kg. Let's check our data.\
If the record was not broken - replace incorrect value manually or to NA.
```{r echo=FALSE}
all_participants %>% filter(ID %in% (all_participants %>% 
                                     filter(Weight > 445) %>%  
                                     .[,1,1]))
```

```{r}
all_participants <- all_participants %>% 
  mutate(Weight = ifelse(Weight > 445, NA, Weight))
```


### 'Games' column
We will find problems in seasons naming:

```{r echo=FALSE}
all_participants %>% 
  filter(!(grepl('Summer', Games, fixed = TRUE) | 
         grepl('Winter', Games, fixed = TRUE) |
         is.na(Games)))
```

Fix it:
```{r}
all_participants <- all_participants %>% 
  mutate(Games=sub('S+.*$', 'Summer', Games)) %>% 
  mutate(Games=sub('W+.*$', 'Winter', Games))
```


### 'Year' column
If we know the Game, we also know the year of its holding.\
Let's fix the empty values.

```{r}
all_participants <- all_participants %>% 
  mutate(Year = ifelse((is.na(Year) & !is.na(Games)), 
                       str_extract(Games,'^[[:digit:]]{4}'), 
                       Year))
```

### 'Season' column
If we know the Game, we also know the Season of its holding.\
Let's fix the empty values.

```{r}
all_participants <- all_participants %>% 
  mutate(Season = ifelse((is.na(Season) & !is.na(Games)),
                         str_extract(Games,'[[:alpha:]]+$'), 
                         Season))
```


### Others columns
All is OK:

```{r echo=TRUE}
all_participants %>% filter(!(Medal %in% c('Gold', 'Silver', 'Bronze', NA)))

```

```{r include=TRUE, echo=FALSE}
all_participants %>% group_by(City) %>%  summarise()

all_participants %>% group_by(Event) %>%  summarise() 


```

## Data standardization

### The Height variable standardization

```{r include=FALSE}
all_participants <- all_participants %>%  
  mutate(Height_z_scores = (Height - 
                              mean(Height, na.rm = TRUE)) / sd(Height, na.rm = TRUE))
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
plot_height <- ggplot(all_participants, aes(Height)) +
  geom_density() +
  theme_bw() +
  labs(title="Height of athletes", x = 'Height, cm')

plot_height_z <- ggplot(all_participants, aes(Height_z_scores)) +
  geom_density() +
  theme_bw() +
  labs(title="Height of athletes", x = 'Standardized Height')

gridExtra::grid.arrange(plot_height, plot_height_z, ncol=2)
```


### The Height variable min-max standardization

```{r include=FALSE}
all_participants <- all_participants %>%  
  mutate(Height_min_max_scaled = (Height - 
                                  min(Height, na.rm = TRUE)) / 
                                 (max(Height, na.rm = TRUE) - 
                                  min(Height, na.rm = TRUE)))
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
plot_height_minmax <- ggplot(all_participants, aes(Height_min_max_scaled)) +
  geom_density() +
  theme_bw() +
  labs(title="Height of athletes", x = '"Min-max" standardized Height')

gridExtra::grid.arrange(plot_height, plot_height_minmax, ncol=2)

```



# Olympic statistics

## Youngest at the 1992 Olympics

```{r include=FALSE}
youngest_men_1992 <- all_participants %>% filter(Year == 1992 & Sex == 'M') %>% 
  filter(Age == min(Age, na.rm = TRUE))

youngest_woman_1992 <- all_participants %>% filter(Year == 1992 & Sex == 'F') %>% 
  filter(Age == min(Age, na.rm = TRUE)) 
```

Age of the youngest athletes at the 1992 Olympics is **`r youngest_men_1992[1,4,1]`** years for men, and **`r youngest_woman_1992[1,4,1]`** years for women.

## Mean and sd of Height variable 
```{r include=FALSE}
mean_man <- all_participants %>% filter(Sex == 'M') %>% 
  summarise(mean(Height, na.rm = TRUE))
mean_woman <- all_participants %>% filter(Sex == 'F') %>% 
  summarise(mean(Height, na.rm = TRUE))

sd_woman <- all_participants %>% filter(Sex == 'F') %>% 
  summarise(sd(Height, na.rm = TRUE))
sd_man <- all_participants %>% filter(Sex == 'M') %>% 
  summarise(sd(Height, na.rm = TRUE))

```

For men Height variable has **mean = `r round(mean_man[1,1,1], 2)`** with **standard deviation = `r round(sd_man[1,1,1], 2)`**.

For women Height variable has **mean = `r round(mean_woman[1,1,1], 2)`** with **standard deviation = `r round(sd_woman[1,1,1], 2)`**.

### Mean and standard deviation of Height variable for female participants at the 2000 Games

```{r include=FALSE}
mean_woman_2000 <- all_participants %>% filter(Sex == 'F', Year == 2000) %>% 
  summarise(mean(Height, na.rm = TRUE))

sd_woman_2000 <- all_participants %>% filter(Sex == 'F', Year == 2000) %>% 
  summarise(sd(Height, na.rm = TRUE))
```

For women Height variable has **mean = `r sprintf("%.1f", round(mean_woman_2000[1,1,1], 1))`** with **standard deviation = `r sprintf("%.1f", round(sd_woman_2000[1,1,1], 1))`** at the 2000 Olympics.

## Heaviest athlete in the 2006 Olympics

```{r include=FALSE}
heaviest_2006 <- all_participants %>% 
  filter(Year == 2006) %>% 
  filter(Weight == max(Weight, na.rm = TRUE))
```

The heaviest athlete in the 2006 Olympics participated in the **`r heaviest_2006 %>% select(Sport) %>% .[1,1,1]`** and had weight `r heaviest_2006 %>% select(Weight) %>% .[1,1,1]` kg.

## Gold medals received by women from 1980 to 2010

```{r include=FALSE}
golds_1980_2010 <- all_participants %>% 
  filter(Sex == 'F' & (Year >= 1980 & Year <= 2010) & Medal == 'Gold') 
```

Women received **`r golds_1980_2010 %>% summarise(n()) %>% .[1,1,1]` gold medals** from 1980 to 2010.

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(golds_1980_2010, aes(y = Games)) +
  geom_histogram(stat="count", alpha = 0.3, fill='Blue') +
  geom_text(aes(label = ..count.., fontface = "bold"), stat = "count", position = position_stack(vjust = 0.5)) +
  labs(x='Number of medals',
       y='Games',
       title="Number of Gold medals received by women from 1980 to 2010") + 
  theme_set(theme_bw(base_size = 15))


```

## Number of participation times of athlete John Aalberg in the Olympic Games

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
john_games <- all_participants %>% filter(grepl('John Aalberg', Name))
john_games %>% group_by(Year) %>% select(Name, Games) %>% mutate(N = n())%>% unique()
```


John Aalberg participated in Olympic games `r john_games %>% summarise(n()) %>% .[1,1,1]` times, `r john_games %>% filter(Games == '1992 Winter')%>% summarise(n()) %>% .[1,1,1]` times in the 1992 Winter Games and `r john_games %>% filter(Games == '1994 Winter')%>% summarise(n()) %>% .[1,1,1]` times in the 1994 Winter Games.

## Age groups of participants in the 2008 Olympic Games

The following boundaries were used to divide the participants into groups:\
- group 0: Age < 15;\
- group 1: Age 15-24;\
- group 2: Age 25-34;\
- group 3: Age 35-44;\
- group 4: Age 45-55;\
- group 5: Age > 55.

Each participant was counted only once, according to their ID number.

```{r include=FALSE}
age_groups_2006 <- all_participants %>% 
  filter(Year == 2008) %>% 
  select(ID, Sex, Age) %>% 
  distinct() %>% 
  mutate(Age_group = NA) %>% 
  mutate(Age_group = ifelse(Age < 15, '14 and less', Age_group)) %>% 
  mutate(Age_group = ifelse(Age > 14 & Age < 25, '15-24', Age_group)) %>% 
  mutate(Age_group = ifelse(Age > 24 & Age < 35, '25-34', Age_group)) %>% 
  mutate(Age_group = ifelse(Age > 34 & Age < 45, '35-44', Age_group)) %>% 
  mutate(Age_group = ifelse(Age > 44 & Age < 56, '45-55', Age_group)) %>% 
  mutate(Age_group = ifelse(Age > 55, '56 and more', Age_group)) %>% 
  filter(!is.na(Age_group))
```

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
ggplot(age_groups_2006, aes(y = Age_group)) +
  geom_histogram(stat="count", alpha = 0.3, fill = 'Orange') +
  geom_text(aes(label = ..count.., fontface = "bold"), stat = "count", position = position_stack(vjust = 0.5)) +
  labs(x='Number of participants',
       y='Age group',
       title="Number of participants in different age groups") + 
  theme_set(theme_bw(base_size = 15))
```

Among the participants from 15 to 55 years old, the smallest group is the `r age_groups_2006 %>% filter(Age_group != '14 and less' & Age_group != '56 and more') %>% group_by(Age_group) %>% summarise(Group_size = n()) %>% filter(Group_size == min(Group_size)) %>% .[1,1,1]` age group, the most numerous is the `r age_groups_2006 %>% filter(Age_group != '14 and less' & Age_group != '56 and more') %>% group_by(Age_group) %>% summarise(Group_size = n()) %>% filter(Group_size == max(Group_size)) %>% .[1,1,1]` age group.

## Сhanging of the number of sports in the 2002 Olympics from the 1994 Olympic


The Olympic Games in 1994 and 2002 are the Winter Olympics. From 1994 to 2002, the number of Olympic winter sports increased.

```{r echo=FALSE, message=FALSE, warning=FALSE}
sports_from_94_to_02 <- all_participants %>% 
  filter(Year >= 1994 & Year <= 2002) %>% 
  filter(Season == 'Winter') %>% 
  select(Year, Season, Sport) %>% 
  distinct()

ggplot(sports_from_94_to_02, aes(y = Year)) +
  geom_histogram(stat="count", alpha = 0.3, fill = 'Red') +
  geom_text(aes(label = ..count.., fontface = "bold"), stat = "count", position = position_stack(vjust = 0.5)) +
  labs(x='Number of sports',
       y='Year',
       title="Number of sports at the Winter Olympics 1994-2002") + 
  theme_set(theme_bw(base_size = 15))

sports_answer <- sports_from_94_to_02 %>% 
  filter(Year == 2002 | Year == 1994) %>% 
  group_by(Year)%>% summarise(n = n())

min_year_sports <- sports_answer %>% filter(n == min(n)) %>% .[1,1,1]

sports_answer <- sports_answer %>% 
  mutate(direction = if_else(min_year_sports == 1994, 'increased', 'decreased'))

```

The number of sports in 1994 was `r sports_answer %>% filter(Year == 1994) %>% .[1,2,1]`. The number of sports in 2002 was `r sports_answer %>% filter(Year == 2002) %>% .[1,2,1]`.

The number of sports at the 2002 Olympics, compared to 1994, has **`r sports_answer[1,3,1]` by `r max(sports_answer[,2,]) - min(sports_answer[,2,])` sports, or `r round((max(sports_answer[,2,]) / min(sports_answer[,2,]) - 1) * 100, 2)`%**.


## Top countries for each type of medal for the winter and summer Olympiads separately 
### Winter Games

```{r include=FALSE}
winter_medals <- all_participants %>% 
  mutate(Team = ifelse(Team == 'Unified Team', 'Soviet Union', Team)) %>% 
  filter(Season == 'Winter' & !is.na(Medal)) %>% 
  mutate(Team = sub('-{1}[[:digit:]]{1}$', '', Team))
```

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
winter_medals %>% 
  group_by(Team) %>% 
  summarise(n = n()) %>% 
  filter(n > 20) %>% 
  ggplot(aes(y = reorder(Team, desc(n)), x = n)) +
  geom_histogram(stat="identity", alpha = 0.3, fill = 'Blue') +
  geom_text(aes(label = n, fontface = "bold"), position = position_stack(vjust = 0.5)) +
  labs(x='Number of medals  (all types) in different countries',
       y='Сountry',
       title="Medals in different countries (if more than 20)",
       subtitle = 'Winter Games. Country name when receiving') + 
  theme_set(theme_bw())

```


Top 3 countries with the most Gold medals in Winter games:
```{r echo=FALSE, message=FALSE, warning=FALSE}
winter_medals %>% 
  filter(Medal == 'Gold') %>% 
  group_by(Team) %>% 
  summarise(Gold_Medals = n()) %>% 
  arrange(-Gold_Medals) %>% 
  top_n(n=3)
```

Top 3 countries with the most Silver medals in Winter games:
```{r echo=FALSE, message=FALSE, warning=FALSE}
winter_medals %>% 
  filter(Medal == 'Silver') %>% 
  group_by(Team) %>% 
  summarise(Silver_Medals = n()) %>%
  arrange(-Silver_Medals) %>% 
  top_n(n=3)
```

Top 3 countries with the most Bronze medals in Winter games:
```{r echo=FALSE, message=FALSE, warning=FALSE}
winter_medals %>% 
  filter(Medal == 'Bronze') %>% 
  group_by(Team) %>% 
  summarise(Bronze_Medals = n()) %>%
  arrange(-Bronze_Medals) %>% 
  top_n(n=3)
```

### Summer Games


```{r include=FALSE}
summer_medals <- all_participants %>% 
  mutate(Team = ifelse(Team == 'Unified Team', 'Soviet Union', Team)) %>% 
  filter(Season == 'Summer' & !is.na(Medal)) %>% 
  mutate(Team = sub('-{1}[[:digit:]]{1}$', '', Team))
```

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
summer_medals %>% 
  group_by(Team) %>% 
  summarise(n = n()) %>% 
  filter(n > 200) %>% 
  ggplot(aes(y = reorder(Team, desc(n)), x = n)) +
  geom_histogram(stat="identity", alpha = 0.3, fill = 'Blue') +
  geom_text(aes(label = n, fontface = "bold"), position = position_stack(vjust = 0.5)) +
  labs(x='Number of medals (all types) in different countries',
       y='Сountry',
       title="Medals in different countries (if more than 200)",
       subtitle = 'Summer Games. Country name when receiving') + 
  theme_set(theme_bw())

```

*Medals won by the Unified Team in the 1992 Olympics counted as medals of the Soviet union*.

Top 3 countries with the most Gold medals in Winter games:
```{r echo=FALSE, message=FALSE, warning=FALSE}
summer_medals %>% 
  filter(Medal == 'Gold') %>% 
  group_by(Team) %>% 
  summarise(Gold_Medals = n()) %>% 
  arrange(-Gold_Medals) %>% 
  top_n(n=3)
```

Top 3 countries with the most Silver medals in Winter games:
```{r echo=FALSE, message=FALSE, warning=FALSE}
summer_medals %>% 
  filter(Medal == 'Silver') %>% 
  group_by(Team) %>% 
  summarise(Silver_Medals = n()) %>%
  arrange(-Silver_Medals) %>% 
  top_n(n=3)
```

Top 3 countries with the most Bronze medals in Winter games:
```{r echo=FALSE, message=FALSE, warning=FALSE}
summer_medals %>% 
  filter(Medal == 'Bronze') %>% 
  group_by(Team) %>% 
  summarise(Bronze_Medals = n()) %>%
  arrange(-Bronze_Medals) %>% 
  top_n(n=3)
```

## Comparsion of the height, weight and ages of men and women who competed in the Winter Olympics

Sample size large enough to treat the distribution of height, weight and age as normally distributed values.

```{r message=FALSE, warning=FALSE, include=FALSE}
task_14 <- all_participants %>% 
  filter(Season == 'Winter') %>% 
  select(Height, Weight, Age, Sex)
```


### Height comparison

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(task_14, aes(Height, color=Sex)) +
  geom_density() +
  scale_color_hue(labels = c("female", "male")) +
  theme_bw() +
  labs(title="Height of athletes")
```

Welch Two Sample t-test (two sided) was performed. Alternative hypothesis - the athletes height depends on the sex. 

```{r message=FALSE, warning=FALSE, include=FALSE}
t_test_height <- t.test(Height~Sex, data = task_14)
```

The value of the t-statistic is **`r round(t_test_height$statistic, 3)`** with **`r round(t_test_height$parameter, 1)`** degrees of freedom. \
Welch Two Sample t-test showed that there was a significant difference between the male athletes height compared with female athletes height  (p-value = **`r t_test_height$p.value` < 2.2 10^-16 **).

### Weight comparison


```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(task_14, aes(Weight, color=Sex)) +
  geom_density() +
  scale_color_hue(labels = c("female", "male")) +
  theme_bw() +
  labs(title="Weight of athletes")
```

Welch Two Sample t-test (two sided) was performed. Alternative hypothesis - the athletes weight depends on the sex. 

```{r message=FALSE, warning=FALSE, include=FALSE}
t_test_weight <- t.test(Weight~Sex, data = task_14)
```

The value of the t-statistic is **`r round(t_test_weight$statistic, 3)`** with **`r round(t_test_weight$parameter, 1)`** degrees of freedom. \
Welch Two Sample t-test showed that there was a significant difference between the male athletes weight compared with female athletes weight  (p-value = **`r t_test_weight$p.value` < 2.2 10^-16 **).

### Age comparison


```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(task_14, aes(Age, color=Sex)) +
  geom_density() +
  scale_color_hue(labels = c("female", "male")) +
  theme_bw() +
  labs(title="Age of athletes")
```

Welch Two Sample t-test (two sided) was performed. Alternative hypothesis - the athletes age depends on the sex. 

```{r message=FALSE, warning=FALSE, include=FALSE}
t_test_age <- t.test(Age~Sex, data = task_14)
```

The value of the t-statistic is **`r round(t_test_age$statistic, 3)`** with **`r round(t_test_age$parameter, 1)`** degrees of freedom. \
Welch Two Sample t-test showed that there was a significant difference between the male athletes age compared with female athletes age  (p-value = **`r t_test_age$p.value` < 2.2 10^-16 **).



